---
title: "Data analysis for: The effect of orientation instructions on the recall and reuse of route and survey elements in wayfinding descriptions"
author: "Jakub Krukar (krukar@uni-muenster.de / krukar.staff.ifgi.de)"
output: github_document
always_allow_html: true
---

Krukar, J., Anacta, V. J., & Schwering, A. (2020). The effect of orientation instructions on the recall and reuse of route and survey elements in wayfinding descriptions. *Journal of Environmental Psychology, 68, 101407*. <https://doi.org/10.1016/j.jenvp.2020.101407>

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(readxl)
library(here)
#library(irr)
library(kableExtra)
#library(tidyr)
library(lmerTest)
library(lme4)
library(sjPlot)
#library(piecewiseSEM)
library(ggplot2)
library(ordinal)
library(dunn.test)
source(here("scripts", "1-analysis.R"))
```

The document lists citations from the paper that refer to some statistical output, and follows the citation with a code chunk that calculated this result. Section, table, and figure numbers are identical as those in the paper.

# 5. Data analysis

## 5.4 Individual differences questionnaires

> Because most of the results of these two individual diﬀerences questionnaires signiﬁcantly correlated in our sample (Mental Rotation Test and global self-conﬁdence scale of the Questionnaire on Spatial Strategies: r = .24, p = .028; Mental Rotation Test and the preference for an allocentric reference frame scale of the Questionnaire on Spatial Strategies: r = .23, p = .034; Mental Rotation Test and the preference for the knowledge of cardinal directions scale of the Questionnaire on Spatial Strategies: r = .12, p = .289), we do not use them simultaneously as predictors in our models in order to avoid the issue of multicollinearity.

```{r}
x <- df.sr %>%
  group_by(part.ID) %>%
  summarise(
    MRT = unique(MRT1SCORE),
    scale1 = unique(GlobalSelfConfidence),
    scale2 = unique(SurveyStrtegy),
    scale3 = unique(CardinalDirKnow)
  )
cor.test(x$MRT, x$scale1)
cor.test(x$MRT, x$scale2)
cor.test(x$MRT, x$scale3)
```

# 6 Results

## 6.1 Quantity of route- and survey-related information

### 6.1.1 Visual mode

> In the visual mode, (...) Participants receiving turn-by-turn instructions, on average, included 2.9 local landmarks in their sketch maps; those in the spatial chunking condition reused an average of 6.9 local landmarks; and those who saw orientation instructions included 9 local landmarks on average

Average number of Local Landmarks per instruction type:

```{r message=FALSE, warning=FALSE}
vis %>%
  filter(ldnm.type != "GL") %>%
  group_by(instr.type, part.ID) %>%
  summarise(no.ldnm = sum(included)) %>%
  group_by(instr.type) %>%
  summarise(avg.no.ldnm = mean(no.ldnm))
```

> ...plus an additional 2.2 global landmarks on average.

Average number of Global Landmarks per instruction type:

```{r}
vis %>%
  filter(ldnm.type == "GL") %>%
  group_by(instr.type, part.ID) %>%
  summarise(no.ldnm = sum(included)) %>%
  group_by(instr.type) %>%
  summarise(avg.no.ldnm = mean(no.ldnm))
```

> To test the global landmark recall rate, we conducted a one-sample t-test, comparing the number of global landmarks recalled by participants in the orientation instructions condition (M = 2.2, SD = 0.6), to number 0. This test was signiﬁcant t(27) = 18.84, p \< .001.

```{r}
xyz1 <- vis %>%
  filter(ldnm.type == "GL") %>%
  group_by(instr.type, part.ID) %>%
  summarise(no.ldnm = sum(included)) %>%
  group_by(instr.type)

t.test(xyz1$no.ldnm, mu = 0, alternative = "greater")
```

> Alternative models including other measures of individual diﬀerences were tested, but only the global self-conﬁdence scale signiﬁcantly improved the model's ﬁt compared to the null model (χ2(1) = 8.26, p = .004).

```{r}
# prepare data for models
vis2 <- vis %>% filter(ldnm.type != "GL")
vis2$landmark <- as.character(vis2$landmark)
vis2$landmark <- str_replace(vis2$landmark, "\\(GasStation\\(Right\\).SCVis", "Gas")
vis2$landmark <- str_replace(vis2$landmark, "Café", "Cafe")
vis2$landmark <- word(vis2$landmark, 1, sep = "\\(")
vis2$landmark <- word(vis2$landmark, 1, sep = "\\.")
vis2$landmark <- as.factor(vis2$landmark)


vis40 <- glmer(included ~ instr.type +
  (1 | part.ID) + (1 + instr.type | landmark), data = vis2, family = binomial("logit"), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
vis41 <- glmer(included ~ instr.type + GlobalSelfConfidence +
  (1 | part.ID) + (1 + instr.type | landmark), data = vis2, family = binomial("logit"), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
vis42 <- glmer(included ~ instr.type + SurveyStrtegy +
  (1 | part.ID) + (1 + instr.type | landmark), data = vis2, family = binomial("logit"), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
vis43 <- glmer(included ~ instr.type + CardinalDirKnow +
  (1 | part.ID) + (1 + instr.type | landmark), data = vis2, family = binomial("logit"), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
vis44 <- glmer(included ~ instr.type + MRT1SCORE +
  (1 | part.ID) + (1 + instr.type | landmark), data = vis2, family = binomial("logit"), control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

anova(vis40, vis41, vis42, vis43, vis44)
```

> Table 3. A logit mixed model describing the impact of instruction type and the global self-conﬁdence scale questionnaire score on the odds of including a landmark in the sketch map, in the visual mode.

```{r  message=FALSE, warning=FALSE}
tab_model(vis41)
```

### 6.1.2 Verbal mode

> In the verbal mode, (...) Spatial chunking participants included on average 4.9 (out of 5) local landmarks, and the orientation instructions participants included on average 8.9 (out of 11) local landmarks

Average number of Local Landmarks per instruction type:

```{r message=FALSE, warning=FALSE}
ver %>%
  filter(ldnm.type != "GL") %>%
  group_by(instr.type, part.ID) %>%
  summarise(no.ldnm = sum(included)) %>%
  group_by(instr.type) %>%
  summarise(avg.no.ldnm = mean(no.ldnm))
```

> ...plus an additional 1.9 global landmarks (out of 2), on average.

Average number of Global Landmarks per instruction type:

```{r}
ver %>%
  filter(ldnm.type == "GL") %>%
  group_by(instr.type, part.ID) %>%
  summarise(no.ldnm = sum(included)) %>%
  group_by(instr.type) %>%
  summarise(avg.no.ldnm = mean(no.ldnm))
```

> To test the global landmark recall rate, we conducted a one-sample, one-tailed t-test, comparing the number of global landmarks recalled by participants in the orientation instructions condition (M = 1.9, SD = 0.3) to number 0. This test was signiﬁcant t(27) = 31.80, p \< .001.

```{r}
xyz2 <- ver %>%
  filter(ldnm.type == "GL") %>%
  group_by(instr.type, part.ID) %>%
  summarise(no.ldnm = sum(included)) %>%
  group_by(instr.type)

t.test(xyz2$no.ldnm, mu = 0, alternative = "greater")
```

> Furthermore, the Mental Rotation Test score was the only individual diﬀerence, including which signiﬁcantly improved the model's ﬁt compared to the null model (χ2(1) = 4.58, p = .032).

```{r}
ver2 <- ver %>% filter(ldnm.type != "GL")

ver40 <- glmer(included ~ instr.type +
  (1 | part.ID) + (1 | instr.type:landmark), data = ver2, family = binomial("logit"))
ver41 <- glmer(included ~ instr.type + MRT1SCORE +
  (1 | part.ID) + (1 | instr.type:landmark), data = ver2, family = binomial("logit"))
ver42 <- glmer(included ~ instr.type + GlobalSelfConfidence +
  (1 | part.ID) + (1 | instr.type:landmark), data = ver2, family = binomial("logit"))
ver43 <- glmer(included ~ instr.type + SurveyStrtegy +
  (1 | part.ID) + (1 | instr.type:landmark), data = ver2, family = binomial("logit"))
ver44 <- glmer(included ~ instr.type + CardinalDirKnow +
  (1 | part.ID) + (1 | instr.type:landmark), data = ver2, family = binomial("logit"))
anova(ver40, ver41, ver42, ver43, ver44)
```

> Table 4. A logit mixed model describing the impact of instruction type and the Mental Rotation Test score on the odds of including a landmark in the sketch map, in the verbal mode.

```{r message=FALSE, warning=FALSE}
tab_model(ver41)
```

## 6.2 Quality of sketch maps

### 6.2.1 Visual

> Table 5. Number of sketch maps that received each of the possible scores (0–3) from the independent raters, in each condition, in the visual mode. The condition and mode from which each sketch map originated was unknown to the raters.

```{r  message=FALSE, warning=FALSE}
final_mapscore <- read_excel(here("data", "final_mapscore.xlsx"), sheet = "Tabellenblatt1")

final_mapscore <- final_mapscore %>% 
  select(map_score, part.ID = participant, instr.media = mode, instr.type = cond)

final_mapscore$part.ID <- str_replace(final_mapscore$part.ID, fixed(" "), "")

df.sr.ver <- left_join(df.sr.ver, final_mapscore)

df.sr.ver$instr.type <- recode(df.sr.ver$instr.type, OI = "Orientation Information", SC = "Spatial Chunking", TbT = "Turn-by-turn")

df.sr.vis <- left_join(df.sr.vis, final_mapscore)

df.sr.vis$instr.type <- recode(df.sr.vis$instr.type, OI = "Orientation Information", SC = "Spatial Chunking", TbT = "Turn-by-turn")

df.sr.vis %>%
  group_by(instr.type) %>%
  count(map_score) %>%
  spread(key = map_score, value = n, fill = 0) %>%
  rename(`condition` = `instr.type`, `0:nonsense` = `0`, `1:bad` = `1`, `2:medium` = `2`, `3:good` = `3`) %>%
  kable() %>%
  kable_styling()
```

> Including global self-conﬁdence scale improved the model's ﬁt best, compared to the null model (LR(1) = 9.28, p = .002).

```{r}
# prepare data for models
df.sr.vis <- left_join(
  df.sr.vis,
  df.sr[
    df.sr$instr.media == "visual",
    c("part.ID", "GlobalSelfConfidence", "SurveyStrtegy", "CardinalDirKnow", "MRT1SCORE", "MRT2SCORE")
  ]
)

df.sr.vis$map_score <- as.ordered(df.sr.vis$map_score)

df.sr.vis <- df.sr.vis %>% mutate_at(1:3, factor)

mapscore.visual0 <- clm(map_score ~ instr.type, data = df.sr.vis)
mapscore.visual1 <- clm(map_score ~ instr.type + MRT1SCORE, data = df.sr.vis)
mapscore.visual2 <- clm(map_score ~ instr.type + GlobalSelfConfidence, data = df.sr.vis)
mapscore.visual3 <- clm(map_score ~ instr.type + SurveyStrtegy, data = df.sr.vis)
mapscore.visual4 <- clm(map_score ~ instr.type + CardinalDirKnow, data = df.sr.vis)

anova(mapscore.visual0, mapscore.visual1, mapscore.visual2, mapscore.visual3, mapscore.visual4)
anova(mapscore.visual0, mapscore.visual2)
```

> Table 6. A proportional odds models for ordinal data demonstrating the inﬂuence of condition on the sketch map's quality score, in the visual mode.

```{r  echo=FALSE, message=FALSE, warning=FALSE}
tab_model(mapscore.visual2)
```

### 6.2.2 Verbal

> Table 7. Number of sketch maps that received each of the possible scores (0–3) from the independent raters, in each condition, in the verbal mode. The condition and mode from which each sketch map originated was unknown to the raters.

```{r message=FALSE, warning=FALSE}
df.sr.ver %>%
  group_by(instr.type) %>%
  count(map_score) %>%
  spread(key = map_score, value = n, fill = 0) %>%
  rename(`condition` = `instr.type`, `0:nonsense` = `0`, `1:bad` = `1`, `2:medium` = `2`, `3:good` = `3`) %>%
  kable() %>%
  kable_styling()
```

> Including the results of the Mental Rotation Test improved the model's ﬁt, compared to the null model (LR(1) = 12.39, p \< .001).

```{r}
# prepare data for models
df.sr.ver <- left_join(
  df.sr.ver,
  df.sr[
    df.sr$instr.media == "verbal",
    c("part.ID", "GlobalSelfConfidence", "SurveyStrtegy", "CardinalDirKnow", "MRT1SCORE", "MRT2SCORE")
  ]
)

df.sr.ver$map_score <- as.ordered(df.sr.ver$map_score)

df.sr.ver <- df.sr.ver %>% mutate_at(1:3, factor)

mapscore.verbal0 <- clm(map_score ~ instr.type, data = df.sr.ver)
mapscore.verbal1 <- clm(map_score ~ instr.type + MRT1SCORE, data = df.sr.ver)
mapscore.verbal2 <- clm(map_score ~ instr.type + GlobalSelfConfidence, data = df.sr.ver)
mapscore.verbal3 <- clm(map_score ~ instr.type + SurveyStrtegy, data = df.sr.ver)
mapscore.verbal4 <- clm(map_score ~ instr.type + CardinalDirKnow, data = df.sr.ver)
anova(mapscore.verbal0, mapscore.verbal1, mapscore.verbal2, mapscore.verbal3, mapscore.verbal4)
```

> Table 8. A proportional odds models for ordinal data demonstrating the inﬂuence of condition on the sketch map's quality score, in the verbal mode.

```{r echo=FALSE, warning=FALSE, message=FALSE}
tab_model(mapscore.verbal1)
```

## 6.3 Type of sketch maps: survey-likeness and route-likeness

> Fig. 6. Survey-likeness and route-likeness scores of sketches from all experimental conditions. Each symbol represents a single sketch map. The scales were ordinal all symbols that are adjacent to each other represent the same number on both scales.

*Note: this table was manually post-processed in graphical editing software. The automatically generated figure plotted multiple objects on top each other (e.g., only one red circle is visible on position 5,5 in the "verbal" panel, while in fact these are multiple circles plotted on top of each other. They were moved apart manually).*

```{r  echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = df.sr, aes(x = surveyness, y = routeness, shape = instr.type, colour = instr.type)) +
  geom_point(size = 5) +
  geom_hline(data = dfs, aes(yintercept = mean.routeness), linetype = "dotted") +
  geom_vline(data = dfs, aes(xintercept = mean.surveyness), linetype = "dotted") +
  scale_y_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6), limits = c(-0.5, 6.5)) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 6), limits = c(-0.5, 6.5)) +
  facet_grid(. ~ instr.media) +
  theme_bw()
```

### 6.3.1 Visual

> Table 9. Number of sketch maps that received each raw score of the route-likeness and of the survey-likeness scale, in the visual mode.

```{r echo=FALSE, message=FALSE, warning=FALSE}
vis.routelikeness <- df.sr.vis %>%
  group_by(instr.type) %>%
  count(routeness) %>%
  spread(key = routeness, value = n, fill = 0) %>%
  rename(` ` = `instr.type`) %>%
  add_column(`0` = 0) %>%
  select(1, `0`, everything())

vis.surveylikeness <- df.sr.vis %>%
  group_by(instr.type) %>%
  count(surveyness) %>%
  spread(key = surveyness, value = n, fill = 0) %>%
  rename(` ` = `instr.type`) %>%
  add_column(`6` = 0)

bind_rows(vis.routelikeness, vis.surveylikeness) %>%
  kable() %>%
  kable_styling() %>%
  pack_rows("Route-likeness", 1, 3) %>%
  pack_rows("Survey-likeness", 4, 6)
```

> We present the models with the single measure of individual differences which best improved the model's ﬁt. For the route-likeness variable, this was the Mental Rotation Test score (LR(1) = 11.50, p \< .001 compared to the corresponding null model);

```{r}
# prepare data for models
df.sr.ord.vis <- filter(df.sr.ord, instr.media == "visual")
df.sr.ord.ver <- filter(df.sr.ord, instr.media == "verbal")

r.vis0 <- clm(routeness ~ instr.type, link = "logit", data = df.sr.ord.vis)
r.vis1 <- clm(routeness ~ instr.type + MRT1SCORE, link = "logit", data = df.sr.ord.vis)
r.vis2 <- clm(routeness ~ instr.type + GlobalSelfConfidence, link = "logit", data = df.sr.ord.vis)
r.vis3 <- clm(routeness ~ instr.type + SurveyStrtegy, link = "logit", data = df.sr.ord.vis)
r.vis4 <- clm(routeness ~ instr.type + CardinalDirKnow, link = "logit", data = df.sr.ord.vis)
anova(r.vis0, r.vis1, r.vis2, r.vis3, r.vis4)
```

> ...for the survey-likeness variable, this was the global self-conﬁdence scale of the Questionnaire on Spatial Strategies (LR(1) = 5.47, p = .019 compared to the corresponding null model).

```{r}
s.vis0 <- clm(surveyness ~ instr.type, link = "logit", data = df.sr.ord.vis)
s.vis1 <- clm(surveyness ~ instr.type + MRT1SCORE, link = "logit", data = df.sr.ord.vis)
s.vis2 <- clm(surveyness ~ instr.type + GlobalSelfConfidence, link = "logit", data = df.sr.ord.vis)
s.vis3 <- clm(surveyness ~ instr.type + SurveyStrtegy, link = "logit", data = df.sr.ord.vis)
s.vis4 <- clm(surveyness ~ instr.type + CardinalDirKnow, link = "logit", data = df.sr.ord.vis)
anova(s.vis0, s.vis1, s.vis2, s.vis3, s.vis4)
anova(s.vis0, s.vis2)
```

> Table 10. Two proportional odds models for ordinal data demonstrating the inﬂuence of condition on the sketch map's route-likeness (left) and survey-likeness (right), in the visual mode.

*Note: Figure rows were re-arranged manually.*

```{r message=FALSE, warning=FALSE}
tab_model(r.vis1, s.vis2)
```

### 6.3.2 Verbal

> Table 11. Number of sketch maps that received each raw score of the route-likeness and of the survey-likeness scale, in the verbal mode.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ver.routelikeness <- df.sr.ver %>%
  group_by(instr.type) %>%
  count(routeness) %>%
  spread(key = routeness, value = n, fill = 0) %>%
  rename(` ` = `instr.type`) %>%
  add_column(`0` = 0) %>%
  select(1, `0`, everything())

ver.surveylikeness <- df.sr.ver %>%
  group_by(instr.type) %>%
  count(surveyness) %>%
  spread(key = surveyness, value = n, fill = 0) %>%
  rename(` ` = `instr.type`) %>%
  add_column(`6` = 0)

bind_rows(ver.routelikeness, ver.surveylikeness) %>%
  kable() %>%
  kable_styling() %>%
  pack_rows("Route-likeness", 1, 3) %>%
  pack_rows("Survey-likeness", 4, 6)
```

> Including the results of the Mental Rotation Test signiﬁcantly improved the model's ﬁt, compared to the null model (LR(1) = 4.66, p = .030).

```{r}
r.ver <- clm(routeness ~ instr.type + MRT1SCORE, link = "logit", data = df.sr.ord.ver)
r.ver0 <- clm(routeness ~ instr.type , link = "logit", data = df.sr.ord.ver)
anova(r.ver, r.ver0)
```

> Table 12. A proportional odds model for ordinal data demonstrating the inﬂuence of condition on the sketch map's route-likeness, in the verbal mode.

```{r}
tab_model(r.ver)
```

> We therefore report the result of the Dunn's test for the survey-likeness variable instead: Comparisons between all three groups were statistically signiﬁcant (Dunn's z = 5.03 for the orientation instructions vs spatial chunking comparison; z = 8.40 for the orientation instructions vs turn-by-turn comparison; z = 3.37 for the spatial chunking vs turn-by-turn comparison; all p-values \< 0.001).

```{r}
# REASON for dunn.test:
# responses too similar to be informative

# df.sr.ord.ver2 <- df.sr.ord.ver %>% filter(instr.type != "TbT")
# s.ver    <- clm(surveyness ~ instr.type, link="logit", data=df.sr.ord.ver2)
# s    <- clm(surveyness ~ instr.type+GlobalSelfConfidence, link="logit", data=df.sr.ord.ver)
# https://stats.stackexchange.com/questions/187841/how-to-deal-with-clmm-warning-hessian-is-numerically-singular

dunn <- dunn.test(df.sr.ver$surveyness, df.sr.ver$instr.type)
```
